---
title: "SMMA - Sentinment Analytics"
format: html
editor: visual

execute: 
  error: false
  warnings: false
  message: false
---

## SENTIMENT ANALYTICS

## 1. Library Installation 

```{r}
#install.packages("tidytext")
#install.packages("wordcloud")
#install.packages("RColorBrewer")
#install.packages("scales")

library(tidyverse)      # dplyr, ggplot2, readr, stringr, tidyr
library(tidytext)       # unnest_tokens(), stop_words, get_sentiments()
library(wordcloud)
library(RColorBrewer)
library(scales)
library(readr)
```

```{r}
library(readxl)
raw <- read_excel("C:\\Users\\Sachin\\OneDrive\\Desktop\\LIFE STORE\\MANIPAL PROGRESS\\SEM-3\\SMMA\\sentiment_lab_data.xlsx")

View(raw)# <- EDIT THIS PATH

dim(raw)                 # expect 500 rows x 13 columns
glimpse(raw)
head(raw, 5)
summary(raw)
```

```{r}
# Missing values per column
colSums(is.na(raw))

# Category levels BEFORE cleaning - note the inconsistent spellings
table(raw$platform, useNA = "ifany")
table(raw$verified_account, useNA = "ifany")
table(raw$region, useNA = "ifany")
table(raw$human_label)

```

## 2. Data Cleaning 
### 2A. Standardise the categorical variables

```{r}
dat <- raw %>%
  mutate(
    platform = str_to_lower(str_trim(platform)),
    platform = case_when(
      str_detect(platform, "insta")            ~ "Instagram",
      str_detect(platform, "youtube")          ~ "YouTube",
      str_detect(platform, "^x|twitter")       ~ "X",
      str_detect(platform, "face")             ~ "Facebook",
      TRUE                                     ~ "Other"),
    verified_account = if_else(str_to_lower(str_trim(verified_account)) == "yes",
                               "Yes", "No"),
    region      = na_if(str_trim(region), ""),
    post_date   = as.Date(post_date),
    week        = floor_date(post_date, unit = "week"),
    human_label = factor(human_label, levels = c("Negative", "Neutral", "Positive"))
  )

table(dat$platform)          # now only 4 clean levels

```

### 2B. Duplicate (reposted / spam) comments

```{r}
sum(duplicated(dat$comment_text))            # how many exact repeats?
dat %>% filter(duplicated(comment_text) |
               duplicated(comment_text, fromLast = TRUE)) %>%
  arrange(comment_text) %>%
  select(comment_id, platform, comment_text) %>%
  print(n = 20)

dat <- dat %>% distinct(comment_text, .keep_all = TRUE)   # keep first occurrence
nrow(dat)                                                 # rows retained
```

### 2C. Missing engagement values

```{r}
# Median imputation within platform (report the count you imputed)
dat <- dat %>%
  group_by(platform) %>%
  mutate(likes_imp = if_else(is.na(likes),
                             median(likes, na.rm = TRUE), as.numeric(likes))) %>%
  ungroup()
```

### 2D. Text cleaning

```{r}
dat <- dat %>%
  mutate(
    text_clean = comment_text,
    text_clean = str_replace_all(text_clean, "http[s]?://\\S+", " "),   # URLs
    text_clean = str_replace_all(text_clean, "@\\w+", " "),             # mentions
    text_clean = str_replace_all(text_clean, "#\\w+", " "),             # hashtags
    text_clean = str_replace_all(text_clean, "[^\\x01-\\x7F]", " "),    # emoji / non-ASCII
    text_clean = str_to_lower(text_clean),
    text_clean = str_replace_all(text_clean, "[[:digit:]]+", " "),
    text_clean = str_replace_all(text_clean, "[[:punct:]]", " "),
    text_clean = str_squish(text_clean),                                # collapse spaces
    n_words    = str_count(text_clean, "\\S+")
  )

dat %>% select(comment_id, comment_text, text_clean) %>% head(3) %>% as.data.frame()
summary(dat$n_words)
```

## 3. DESCRIPTIVE PROFILE

```{r}
## ---------------------------------------------------------------------------
dat %>% count(platform, sort = TRUE) %>% mutate(pct = percent(n / sum(n)))
dat %>% count(brand, human_label) %>% pivot_wider(names_from = human_label, values_from = n)

dat %>%
  group_by(human_label) %>%
  summarise(comments  = n(),
            mean_likes = round(mean(likes_imp), 1),
            mean_repl  = round(mean(replies), 1),
            mean_stars = round(mean(star_rating, na.rm = TRUE), 2))

ggplot(dat, aes(x = fct_infreq(platform), fill = human_label)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  labs(title = "Human-coded sentiment mix by platform",
       x = "Platform", y = "Share of comments", fill = "Sentiment")

```
## 4. TOKENISATION AND WORD FREQUENCY

```{r}
## ---------------------------------------------------------------------------
custom_stop <- tibble(word = c("brand", "product", "amp", "just", "really",
                               tolower(unique(dat$brand))))

tokens <- dat %>%
  select(comment_id, platform, brand, human_label, text_clean) %>%
  unnest_tokens(word, text_clean) %>%
  anti_join(stop_words, by = "word") %>%
  anti_join(custom_stop, by = "word")

nrow(tokens)                                   # total tokens after stopword removal

top_words <- tokens %>% count(word, sort = TRUE) %>% slice_head(n = 20)
top_words

ggplot(top_words, aes(x = n, y = fct_reorder(word, n))) +
  geom_col(fill = "steelblue") +
  labs(title = "Twenty most frequent terms", x = "Frequency", y = NULL)
```


```{r}
# Words that distinguish positive from negative comments
tokens %>%
  filter(human_label != "Neutral") %>%
  count(human_label, word) %>%
  group_by(human_label) %>%
  slice_max(n, n = 10) %>%
  ungroup() %>%
  ggplot(aes(n, fct_reorder(word, n), fill = human_label)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ human_label, scales = "free_y") +
  labs(title = "Top terms within each sentiment class", x = "Frequency", y = NULL)

```

```{r}

# Word cloud
set.seed(6306)
with(count(tokens, word, sort = TRUE),
     wordcloud(word, n, max.words = 80, min.freq = 3,
               colors = brewer.pal(8, "Dark2"), random.order = FALSE))
```

## 5. LEXICON-BASED SENTIMENT SCORING (BING)

```{r}
## ---------------------------------------------------------------------------
bing <- get_sentiments("bing")     # bundled with tidytext, works offline

comment_scores <- tokens %>%
  inner_join(bing, by = "word") %>%
  count(comment_id, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(net_score = positive - negative)

dat <- dat %>%
  left_join(comment_scores, by = "comment_id") %>%
  mutate(
    positive  = replace_na(positive, 0),
    negative  = replace_na(negative, 0),
    net_score = replace_na(net_score, 0),
    pred_label = factor(case_when(net_score > 0 ~ "Positive",
                                  net_score < 0 ~ "Negative",
                                  TRUE          ~ "Neutral"),
                        levels = c("Negative", "Neutral", "Positive"))
  )

summary(dat$net_score)
table(dat$pred_label)

ggplot(dat, aes(x = net_score)) +
  geom_histogram(binwidth = 1, fill = "grey30") +
  labs(title = "Distribution of net Bing sentiment score",
       x = "Positive words - negative words", y = "Comments")


```
## 6. VALIDATION AGAINST THE HUMAN-CODED LABEL

```{r}
cm <- table(Predicted = dat$pred_label, Actual = dat$human_label)
cm

accuracy <- sum(diag(cm)) / sum(cm)
round(accuracy, 3)                       # <- report this number in your answer

# Per-class precision, recall and F1
perf <- tibble(class = colnames(cm)) %>%
  rowwise() %>%
  mutate(
    TP        = cm[class, class],
    precision = TP / sum(cm[class, ]),
    recall    = TP / sum(cm[, class]),
    f1        = 2 * precision * recall / (precision + recall)
  ) %>%
  ungroup() %>%
  mutate(across(precision:f1, ~ round(.x, 3)))
perf

# Cohen's kappa (agreement corrected for chance)
p_obs <- accuracy
p_exp <- sum(rowSums(cm) / sum(cm) * colSums(cm) / sum(cm))
kappa <- (p_obs - p_exp) / (1 - p_exp)
round(kappa, 3)

# Inspect the failures - this is where the marks are
dat %>%
  filter(pred_label != human_label) %>%
  select(comment_id, human_label, pred_label, net_score, comment_text) %>%
  slice_head(n = 15) %>%
  as.data.frame()

# The worst kind of error: negative comments called positive
dat %>%
  filter(human_label == "Negative", pred_label == "Positive") %>%
  select(comment_id, net_score, comment_text) %>%
  as.data.frame()


```
## 7. WHY THE LEXICON FAILS: NEGATION AND SARCASM

```{r}
## ---------------------------------------------------------------------------
negators <- c("not", "no", "never", "without", "hardly", "zero", "nothing", "cannot")

bigrams <- dat %>%
  select(comment_id, text_clean) %>%
  unnest_tokens(bigram, text_clean, token = "ngrams", n = 2) %>%
  separate(bigram, c("w1", "w2"), sep = " ")

negated_terms <- bigrams %>%
  filter(w1 %in% negators) %>%
  inner_join(bing, by = c("w2" = "word")) %>%
  count(w1, w2, sentiment, sort = TRUE)
negated_terms                   # every row here was scored with the wrong sign

# Simple correction: flip the sign of any sentiment word preceded by a negator
correction <- bigrams %>%
  filter(w1 %in% negators) %>%
  inner_join(bing, by = c("w2" = "word")) %>%
  mutate(adj = if_else(sentiment == "positive", -2, 2)) %>%
  group_by(comment_id) %>%
  summarise(adjustment = sum(adj), .groups = "drop")

dat <- dat %>%
  left_join(correction, by = "comment_id") %>%
  mutate(adjustment  = replace_na(adjustment, 0),
         score_adj   = net_score + adjustment,
         pred_adj    = factor(case_when(score_adj > 0 ~ "Positive",
                                        score_adj < 0 ~ "Negative",
                                        TRUE          ~ "Neutral"),
                              levels = c("Negative", "Neutral", "Positive")))

cm2 <- table(Predicted = dat$pred_adj, Actual = dat$human_label)
cm2
accuracy2 <- sum(diag(cm2)) / sum(cm2)
round(accuracy2, 3)                      # compare with `accuracy` above
round(accuracy2 - accuracy, 3)           # did negation handling help?
```

## 8. SENTIMENT ACROSS PLATFORMS AND BRANDS

```{r}
## ---------------------------------------------------------------------------
brand_tab <- dat %>%
  count(brand, human_label) %>%
  group_by(brand) %>%
  mutate(share = round(n / sum(n), 3)) %>%
  ungroup()
brand_tab %>% pivot_wider(id_cols = brand, names_from = human_label, values_from = share)

# Net sentiment index per brand: (%positive - %negative)
dat %>%
  group_by(brand) %>%
  summarise(n = n(),
            NSI = round(100 * (mean(human_label == "Positive") -
                               mean(human_label == "Negative")), 1)) %>%
  arrange(desc(NSI))

# Is sentiment independent of platform?
chisq_platform <- chisq.test(table(dat$platform, dat$human_label))
chisq_platform
round(chisq_platform$expected, 1)        # check the >5 assumption before you interpret

ggplot(brand_tab, aes(x = brand, y = share, fill = human_label)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = percent) +
  labs(title = "Sentiment share by brand", x = NULL, y = "Share", fill = "Sentiment")

```
## 9. DOES SENTIMENT DRIVE ENGAGEMENT?

```{r}
## ---------------------------------------------------------------------------
dat <- dat %>% mutate(log_likes = log1p(likes_imp))

dat %>%
  group_by(human_label) %>%
  summarise(n = n(), mean_likes = round(mean(likes_imp), 1),
            median_likes = median(likes_imp), sd = round(sd(likes_imp), 1))

aov_fit <- aov(log_likes ~ human_label, data = dat)
summary(aov_fit)
TukeyHSD(aov_fit)

ggplot(dat, aes(x = human_label, y = likes_imp, fill = human_label)) +
  geom_boxplot(show.legend = FALSE, outlier.alpha = 0.3) +
  scale_y_log10(labels = comma) +
  labs(title = "Likes by sentiment class (log scale)", x = NULL, y = "Likes")

# Control for audience size and account type
model <- lm(log_likes ~ human_label + log1p(user_followers) +
              verified_account + platform, data = dat)
summary(model)

# Replies as a proxy for controversy
lm(log1p(replies) ~ human_label + log1p(user_followers), data = dat) %>% summary()

```
## 10. SENTIMENT OVER TIME

```{r}
## ---------------------------------------------------------------------------
weekly <- dat %>%
  group_by(week) %>%
  summarise(comments = n(),
            NSI = 100 * (mean(human_label == "Positive") -
                         mean(human_label == "Negative")), .groups = "drop")

ggplot(weekly, aes(x = week, y = NSI)) +
  geom_line(colour = "grey40") +
  geom_point(aes(size = comments), colour = "steelblue") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Weekly net sentiment index",
       x = NULL, y = "NSI  (%positive - %negative)", size = "Comments")


```
## 11. EXPORT
```{r}
write_csv(dat %>% select(comment_id, platform, brand, post_date, human_label,
                         pred_label, pred_adj, net_score, score_adj,
                         likes_imp, replies, comment_text),
          "sentiment_scored_output.csv")

sessionInfo()      # paste this into your submission for reproducibility

```

